Analysis of Prenatal Depression and Anxiety during Covid-19

### 2024-03-25 outputs/paper.qmd
finalize the tables, graphs and models

---
title: "Analysis of Prenatal Depression and Anxiety during Covid-19"
author: 
  - Boxuan Yi
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: ""
date: "24 March 2024"
date-format: "D MMMM YYYY"
number-sections: true
thanks: "Code and data in this analysis is available at: https://github.com/Elaineyi1/Prenatal_Mental_Health"
bibliography: references.bib
toc: true
---
# Introduction
```{r}
#| include: false
library(arrow)
library(readr)
library(here)
library(ggplot2)
library(janitor)
library(rstanarm)
library(dplyr)
library(modelsummary)
library(broom)
library(knitr)
```

@@

# Data

```{r}
#| include: false
age_levels <- c("Under 27", "27 to 34", "35 to 39", "40 and above")
income_levels <- c("Less than $20,000", "$20,000 - $39,999", "$40,000 - $69,999", 
                   "$70,000 - $99,999", "$100,000 - $124,999", "$125,000 - $149,999",
                   "$150,000 - $174,999", "$175,000 - $199,999", "$200,000+")
month_levels <- c("Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
clean_data$age_group <- factor(clean_data$age_group, levels = age_levels)
clean_data$household_income <- factor(clean_data$household_income, levels = income_levels)
clean_data$month <- factor(clean_data$month, levels = month_levels)
```

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-income
#| tbl-cap: Levels of Depression and Anxiety Among Different Income Groups
clean_data |> group_by(household_income) |> 
  summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
  rename("Household Income in 2019" = household_income, 
         "Average Level of Depression" = average_depression, 
         "Average Level of Anxiety" = average_anxiety) |>
  kable()
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-age
#| tbl-cap: Levels of Depression and Anxiety Among Different Age Groups
clean_data |> group_by(age_group, language) |> 
  summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
  rename("Age Group" = age_group, 
         "Survey Language" = language,
         "Average Level of Depression" = average_depression, 
         "Average Level of Anxiety" = average_anxiety) |>
  kable()
```

@@

# Model
```{r}
#| include: false
depression_model <- readRDS(file = here::here("model/depression.rds"))
```

@@

# Results
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-depression_model
#| tbl-cap: Estimating Prenatal Depression given Age, Household Income, Month of Delivery and Choice of Language
modelsummary(
  list(
    "Level of Depression" = depression_model))
```

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-depression_interval
#| fig-cap: The Coefficients and Confidence Interval in the Depression Model
modelplot(depression_model, conf_level = 0.95) +
  labs(x = "95% Level of Depression Interval")
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-anxiety_interval
#| fig-cap: The Coefficients and Confidence Interval in the Anxiety Model
modelplot(anxiety_model, conf_level = 0.95) +
  labs(x = "95% Level of Anxiety Interval")
```

```{r}
#| label: fig-danger
#| fig-cap: Association between Perceived Level of Danger and Prenatal Mental Health
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Relationship between Perceived Danger to Self and Depression", "Relationship between Perceived Danger to the Unborn and Depression", "Relationship between Perceived Danger to Self and Anxiety", "Relationship between Perceived Danger to the Unborn and Anxiety"]
clean_data |>
  ggplot(aes(x = threaten_life, y = EPDS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() + 
  labs(x="Perceived Level of Danger", y = "Level of Prenatal Depression") + ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = EPDS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Depression") + ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_life, y = PROMIS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety") + ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = PROMIS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Anxiety") + ylim(0,1)
```

# Discussions
```{r}
#| label: fig-weight_delivery
#| fig-cap: Distribution of Birth Weight and Prenatal Mental Health by Delivery Mode
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Distribution of Birth Weight and Depression by Delivery Mode", "Distribution of Birth Weight and Anxiety by Delivery Mode"]
clean_data |>
  ggplot(aes(x = EPDS, y = birth_weight, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() + labs(x = "The Level of Prenatal Depression", y = "Birth Weight (grams)")

clean_data |>
  ggplot(aes(x = PROMIS, y = birth_weight, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() +
  labs(x = "The Level of Prenatal Anxiety", y = "Birth Weight (grams)")
```

# Appendix
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-danger_baby
#| fig-cap: Distribution of the Preceived Danger to Baby and Preceived Potential Harm to Baby
clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = threaten_baby_harm)) +
  geom_jitter(color='darkgray', alpha=0.4) +
  geom_smooth() +
  theme_classic() 
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-language_income
#| fig-cap: Income Group Distribution Based on the Choice of Survey Language
clean_data |>
  ggplot(aes(x = household_income, fill = language)) +
  stat_count(position = "dodge") +
  theme_minimal() +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

>>>>

---
title: "Analysis of Prenatal Depression and Anxiety during Covid-19"
author: 
  - Boxuan Yi
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: ""
date: "24 March 2024"
date-format: "D MMMM YYYY"
number-sections: true
thanks: "Code and data in this analysis is available at: https://github.com/Elaineyi1/Prenatal_Mental_Health"
bibliography: references.bib
toc: true
---
# Introduction {#sec-introduction}
```{r}
#| include: false
library(arrow)
library(readr)
library(here)
library(ggplot2)
library(janitor)
library(rstanarm)
library(dplyr)
library(modelsummary)
library(broom)
library(knitr)
```

@@

# Data {#sec-data}

```{r}
#| include: false
age_levels <- c("Under 27", "27 to 34", "35 to 39", "40 and above")
income_levels <- c("Less than $20,000", "$20,000 - $39,999", "$40,000 - $69,999", 
                   "$70,000 - $99,999", "$100,000 - $124,999", "$125,000 - $149,999",
                   "$150,000 - $174,999", "$175,000 - $199,999", "$200,000+")
education_levels <- c("Less than high school diploma", "High school diploma", "College/trade school", 
                      "Undergraduate degree", "Masters degree", "Doctoral Degree")
month_levels <- c("Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
clean_data$age_group <- factor(clean_data$age_group, levels = age_levels)
clean_data$household_income <- factor(clean_data$household_income, levels = income_levels)
clean_data$maternal_education <- factor(clean_data$maternal_education, levels = education_levels)
clean_data$month <- factor(clean_data$month, levels = month_levels)
```

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-education_age
#| fig-cap: The Distribution of Age Groups and Level of Education for Pregnant Participants
clean_data |>
  ggplot(mapping = aes(x = maternal_education, fill = age_group)) +
  geom_bar(position = "dodge2") + 
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Individual's Education Level", y = "Number of Observations", fill = "Age Group")
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-age_language
#| tbl-cap: Levels of Depression and Anxiety Among Different Age Groups
clean_data |> group_by(age_group, language) |> 
  summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
  rename("Age Group" = age_group, 
         "Survey Language" = language,
         "Average Level of Depression" = average_depression, 
         "Average Level of Anxiety" = average_anxiety) |>
  kable()
```

```{r}
#| label: fig-average_depression
#| fig-cap: Distribution of Levels of Depression by Education and Household Income
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Distribution of Levels of Depression by Education", "Distribution of Levels of Depression by Household Income"]
clean_data %>%
  group_by(maternal_education) %>%
  summarise(average_depression = mean(EPDS)) %>%
  ggplot(aes(x = maternal_education, y = average_depression)) +
  geom_bar(stat = "identity", fill = "cornflowerblue") + geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.5) +
  labs(x="Education", y="Average Level of Depression")

clean_data %>%
  group_by(household_income) %>%
  summarise(average_depression = mean(EPDS)) %>%
  ggplot(aes(x = household_income, y = average_depression)) + 
  geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.5)+
  geom_bar(stat = "identity", fill = "cornflowerblue") + labs(x="Household Income in 2019", y="Average Level of Depression")
```

```{r}
#| label: fig-average_anxiety
#| fig-cap: Distribution of Levels of Anxiety by Education and Household Income
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Distribution of Levels of Anxiety by Education", "Distribution of Levels of Anxiety by Household Income"]
clean_data %>%
  group_by(maternal_education) %>%
  summarise(average_anxiety = mean(PROMIS)) %>%
  ggplot(aes(x = maternal_education, y = average_anxiety)) +
  geom_bar(stat = "identity", fill = "salmon2") + geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7)+
  labs(x="Education", y="Average Level of Anxiety")

clean_data %>%
  group_by(household_income) %>%
  summarise(average_anxiety = mean(PROMIS)) %>%
  ggplot(aes(x = household_income, y = average_anxiety)) + 
  geom_bar(stat = "identity", fill = "salmon2") +
  geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7) +
  labs(x="Household Income in 2019", y="Average Level of Anxiety")
```

@@

# Model {#sec-model}
```{r}
#| include: false
depression_model <- readRDS(file = here::here("model/depression.rds"))
```

@@

# Results {#sec-results}
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-depression_model
#| tbl-cap: Estimating Prenatal Depression given Age, Household Income, Month of Delivery and Choice of Language
modelsummary(
  list(
    "Level of Depression" = depression_model))
```

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-depression_interval
#| fig-cap: The Coefficients and Confidence Interval in the Depression Model
modelplot(depression_model, conf_level = 0.95, color = 'royalblue3') +
  labs(x = "95% Level of Depression Interval")
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-anxiety_interval
#| fig-cap: The Coefficients and Confidence Interval in the Anxiety Model
modelplot(anxiety_model, conf_level = 0.95, color = 'indianred3') +
  labs(x = "95% Level of Anxiety Interval")
```

# Discussions {#sec-discussions}

```{r}
#| label: fig-weight_delivery
#| fig-cap: Distribution of Birth Weight and Length and Prenatal Mental Health by Delivery Mode
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Distribution of Birth Weight and Depression by Delivery Mode", "Distribution of Birth Weight and Anxiety by Delivery Mode", "Distribution of Birth Length and Depression by Delivery Mode", "Distribution of Birth Length and Anxiety by Delivery Mode"]
clean_data |>
  ggplot(aes(x = EPDS, y = birth_weight, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() + labs(x = "The Level of Prenatal Depression", y = "Birth Weight (grams)")

clean_data |>
  ggplot(aes(x = PROMIS, y = birth_weight, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() +
  labs(x = "The Level of Prenatal Anxiety", y = "Birth Weight (grams)")

clean_data |>
  ggplot(aes(x = EPDS, y = birth_length, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() + labs(x = "The Level of Prenatal Depression", y = "Birth Length (centimeters)")

clean_data |>
  ggplot(aes(x = PROMIS, y = birth_length, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() +
  labs(x = "The Level of Prenatal Anxiety", y = "Birth Length (centimeters)")
```

# Appendix {#sec-appendix}
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-language_income_education
#| fig-cap: Income and Education Groups Distribution Based on the Choice of Survey Language
#| layout-ncol: 2
#| fig-subcap: ["Distribution of Income by Language", "Distribution of Education by Language"]
clean_data |>
  ggplot(aes(x = household_income, fill = language)) +
  stat_count(position = "dodge") +
  theme_minimal() +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") + theme(legend.position = "bottom")

clean_data |>
  ggplot(aes(x = maternal_education, fill = language)) +
  stat_count(position = "dodge") +
  theme_minimal() +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") + theme(legend.position = "bottom")
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-danger_baby
#| fig-cap: Distribution of the Preceived Danger to Baby and Preceived Potential Harm to Baby
clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = threaten_baby_harm)) +
  geom_jitter(color='darkgray', alpha=0.4) +
  geom_smooth() + theme_minimal()
```

```{r}
#| label: fig-danger
#| fig-cap: Association between Perceived Level of Danger and Prenatal Mental Health
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Relationship between Perceived Danger to Self and Depression", "Relationship between Perceived Danger to the Unborn and Depression", "Relationship between Perceived Danger to Self and Anxiety", "Relationship between Perceived Danger to the Unborn and Anxiety"]
clean_data |>
  ggplot(aes(x = threaten_life, y = EPDS, color = language)) +
  geom_smooth(method = "lm") + 
  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
  theme_classic() + 
  labs(x="Perceived Level of Danger", y = "Level of Prenatal Depression")+ ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = EPDS, color = language)) +
  geom_smooth(method = "lm") + 
  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
  theme_classic() +
  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Depression") + ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_life, y = PROMIS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() + geom_hline(yintercept = c(0.4, 0.6), linetype = "dashed", color = "black")+
  labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety")+ ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = PROMIS, color = language)) +
  geom_smooth(method = "lm") + 
  geom_hline(yintercept = c(0.4, 0.6), linetype = "dashed", color = "black") +
  theme_classic() +
  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Anxiety") + ylim(0,1) 
```
